\documentclass[12pt,twoside]{report}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{color}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\usepackage[style=numeric-comp, sorting=none]{biblatex}
\bibliography{xampl}
\addbibresource{references.bib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage[T1]{fontenc}
\usepackage{sourcecodepro}
\lstset{basicstyle=\small\ttfamily}

\title{
    {Phase Transition Continuity Dependence on Edge Evaluation and Acceptance}\\
    {\large Universität Leipzig}
    }
\author{Cameron Perot}
\date{September 2019}


\begin{document}
\maketitle
\pagenumbering{roman}


\begin{center}
\bf{Declaration of Authorship}
\end{center}

I hereby certify that the thesis I am submitting is entirely my own original work except where otherwise indicated. I am aware of the University's regulations concerning plagiarism, including those regulations concerning disciplinary actions that may result from plagiarism. Any use of the works of any other author, in any form, is properly acknowledged at their point of use.
\\\\\\\\\\

\noindent Cameron Perot

\chapter*{Abstract}
When the topic of explosive percolation first came about it was quite controversial in the sense that many people were debating whether the "explosive" phase transitions produced by certain graph evolution processes were continuous or not.
In this thesis I summarize some of the major developments in the field up to this point and go into detail about some of these processes.
Over the years there has been a lot of research into explosive percolation, eventually concluding that algorithms which take only local information into account produce continuous transitions whereas algorithms that make use of global information can produce discontinuous transitions.
Here I also introduce a new model called stochastic edge acceptance (SEA) and analyze it in a framework for proving continuity laid out by previous researchers.
Analysis of the SEA model shows that it exhibits beautiful power law scaling behavior near the critical point, allowing us to determine critical exponents and show that it does indeed produce a continuous transition.


\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

\chapter{Introduction}
\label{ch:introduction}
\input{chapters/introduction.tex}

\chapter{An "Explosive" Transition}
\label{ch:explosive}
\input{chapters/explosive_transition.tex}

\chapter{A New Model}
\label{ch:sea}
\input{chapters/sea.tex}

\chapter{Conclusion}
\label{ch:conclusion}
\input{chapters/conclusion.tex}

\appendix
\chapter{Appendix}
\label{ch:appendix}
\input{chapters/appendix.tex}

\printbibliography

\end{document}
